---
title: "EDA.rmd"
author: "Nick Jeschke"
date: "2025-07-07"
output: html_document
---

Exploratory Data Analysis

```{r}
dim(bp_data_all)
colnames(bp_data_all)

```
160 Columns, 26,757 observations

Year in bp_data$Survey_Year: 1999-2017
1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017

Outcome of interest: bp_data$bp_control_accaha
Binary BP control (<140/<90), according to accaha guidelines



Checking those two variables for missingness, number of cycles
```{r}
sum(is.na(bp_data_all$bp_control_accaha))
sum(is.na(bp_data_all$Survey_Year))

unique(bp_data_all$Survey_Year)

```
No missing values for either bp_control or survey_year, surveys from 1999-2017 available



Investigating general trend in BP control over all
```{r}
freq_table <- table(bp_data_all$Survey_Year, bp_data_all$bp_control_accaha)
prop_table <- prop.table(freq_table, margin = 1)

bpbyyear <- barplot(
  t(prop_table),                    
  beside = FALSE,                   
  col = c("lightblue", "salmon"),   
  legend.text = TRUE,
  args.legend = list(title = "BP Controlled", x = "bottomleft"),
  xlab = "Year",
  ylab = "Proportion",
  main = "Proportion of BP Control by Year"
)

```
Generally decreases from 1999 to 2013, but then starts to increase slightly after that



Checking missingness by row and column
```{r}
missing_props <- colMeans(is.na(bp_data_all))
hist(missing_props, xlab="Column Missingness", main="Proportion of Missing Data by Column")

row_missing_counts <- rowSums(is.na(bp_data_all))
hist(row_missing_counts, xlab = "Row Missingness", main="Proportion of Missing Data by Row")
```
